TODO

### 2015-03-18 ccs.tex
Strip out security updates and android ecosystem stuff from esorics paper and put in a ccs paper and make the esorics paper meet the page limit

\documentclass{llncs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[space]{grffile}
\graphicspath{figures/}
\usepackage{import}
\usepackage{url}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{siunitx}
\usepackage{todonotes}%[disable]
\let\OldTodo\todo
\renewcommand{\todo}{\OldTodo[inline]}
\newcommand{\todolater}[1]{\todo{#1}}%{}%
\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
\addbibresource{securityupdates.bib}
\renewcommand{\bibfont}{\small}

@@

\author{Daniel R.\ Thomas, Alastair R.\ Beresford, Daniel T.\ Wagner and Andrew Rice}

@@

% Blinding function
\newcommand{\identifying}[1]{#1}%{}%
\newcommand{\blindauthors}[1]{#1}%{Paper \#59}%

\begin{document}
\title{Timeliness of security updates for Android}

\author{\blindauthors{Daniel~R.~Thomas \and Daniel~T.~Wagner\and Alastair~R.~Beresford \and Andrew~Rice}}
\institute{\blindauthors{Computer Laboratory,
University of Cambridge,
Cambridge, United Kingdom\\
\email{firstname.lastname@cl.cam.ac.uk}}}

@@

@@

% Hypothesis
% Attempts at fine grained restrictions on running arbitrary code are hampered by updates for security vulnerabilities not reaching users in a timely fashion.

\begin{abstract}
Modern smartphone operating systems protect users from malicious applications by providing a protected execution environment or sandbox.
This approach relies on the absence of active exploits which allow apps to break out of the sandbox.
Many such exploits have been found, and therefore the security of a smartphone operating system relies on the prompt delivery of security updates.
In this paper we tease apart the Android software ecosystem and quantify the extent to which the security sandbox provided by Android protects devices from malicious apps.
We analysed \da\ data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average \daMeanInsecurityPercNominal\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPercNominal\ of devices run the most recent version of Android.
We find there is significant variability in the timely delivery of security updates across different device manufacturers and mobile network operators, and we define a security metric to rank the performance of device manufacturers and network operators in our data set.
\end{abstract}

\section{Introduction}
%Android is the most popular smartphone operating system today with \percMarketShare\ marketshare.

Support for third-party apps is a key feature of modern smartphone operating systems.
Such apps are written by many developers from a wide range of backgrounds, which means neither operating system vendors nor users can fully trust app developers.
Therefore, in order to secure personal data and prevent theft, such as the sending of premium-rate text messages, operating system vendors provide a protected execution environment, or \emph{sandbox}, for apps.

In this paper we quantify the extent to which the app sandbox provided by Android protects devices from malicious apps attempting to gain control over the entire handset.
%We have analysed Android because it is the most popular smartphone operating system by market share (~\percMarketShare) and 
Previous research has shown that in 2012, between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained exploits designed to enable a malicious app to break the sandbox.
What is not known however is the proportion of Android devices which are vulnerable to these exploits over time.
To the best of our knowledge, this is the first paper to provide detailed numbers.

Not all malicious apps need to break the sandbox in order to misbehave.
Indeed, the above figures suggest the majority of malicious apps do not attempt to do so.
We focus on the issue of apps breaking out of the sandbox because malware which is able to take full control of a handset can do significantly more harm and is much harder to remove.
Ordinary malware can be uninstalled by the user, or remotely by Google through the Play Store, and returns the device to a secure state. 
Removing an app which has used an exploit to take control of the handset is unlikely to return the device to secure state.
We describe our threat model and three attack vectors in detail in \S\ref{sec:threatmodel}.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{figures/proportioninsecure}
\caption{Proportion of devices running insecure, maybe secure and secure versions of Android against time.
The red vertical lines are caused by vulnerabilities being discovered with those which have the biggest impact annotated.
This graph is explained further in \S\ref{sec:exp:versionsecurity}.
}
\label{fig:proportioninsecure}
\end{figure}

In order to determine whether an individual smartphone is vulnerable to a particular exploit, we need to combine vulnerability data with the operating system version string and build number from the handset. 
For this we use handset data gathered from \daNumOSDevices\ devices in the \da~\cite{Wagner2013} project.
This analysis allows us to plot the proportion of Android devices in the \da\ data which were running versions of Android known to have vulnerabilities at the time and is shown in Figure~\ref{fig:proportioninsecure}.
The figure shows the proportion of devices split into three categories:
(1)~\emph{secure} because they are running a version of Android with no known vulnerabilities; 
(2)~\emph{insecure} because the device was running an insecure version of Android with a build released before the vulnerability was fixed or discovered; and 
(3)~\emph{maybe secure} because the device is running a version of Android known to be insecure but has received an update which might contain a backported fix as we did not observe that build number until after the vulnerability was known.
Further analysis into the propagation of updates in Android is presented in \S\ref{sec:exp:android_ecosystem}.

@@

In summary, the contributions of this paper are:
\begin{itemize}
 \item We characterise the Android update ecosystem, showing how updates flow between entities, and quantify vulnerabilities by determining what proportion of devices they affected and how that changes over time.
 \item We created an open database to collate quantitative information on vulnerabilities which affect Android, and investigate their time-lines.
 \item We measure the security of Android according to three metrics and compare different device manufacturers and device models to allow device purchasers to differentiate between them based on security.
\end{itemize}

\section{Threat model}
\label{sec:threatmodel}

In this paper we measure whether an attacker can take control of a handset by breaking out of the app sandbox on Android using a known vulnerability.
We consider three attack vectors.
The first attack vector is through the installation of a malicious app on the device.
Android devices support the installation of apps through a number of app marketplaces, email attachments, URLs and via the Android Debug Bridge (ADB).
By default, many Android devices with Google Play installed will prevent the installation of apps from other sources.

Controlling potential sources of app installation is a good security feature because app marketplaces provide additional protection to prevent the listing of malicious apps.
For example, the Google Play marketplace uses Bouncer to automatically analyse apps as well as removing any apps which are reported as malicious.
Nevertheless, Google Play is not immune: 0.02\% of apps are malicious~\cite{Zhou2012a}.
Alternative markets are also popular, particularly in countries such as China where the Google Play is not available. 
Previous work has shown that on alternative markets between 0.20\% and 0.47\% of apps are malicious~\cite{Zhou2012a}.

The second attack vector is to persuade an existing app to download and execute code at runtime.
The most direct method available to an attacker is to upload to a marketplace a seemingly innocent app which contains support for dynamic code loading from a remote server under control of the attacker.
Neither static nor dynamic analysis of this app by the marketplace will uncover any malicious code, since it does exist in the app binary.
Instead the attacker can chose to send a malicious payload directly to the app for execution at a later point in time. Note that dynamic code loading is not actually needed for this to work---it just makes the task easier---since previous work has shown that even on a platform such as iOS, which does not permit dynamic code loading, a ROP-based attack is relatively easy if the attacker creates an app with carefully crafted flaws~\cite{}.

Support for dynamic code loading on Android also supports the third attack vector in which the attacker injects malicious code directly into existing apps already installed on the handset. 
For example, the addJavascriptInterface vulnerability (CVE-2012-6636) allows Javascript running in an Android WebView to execute arbitrary code in the same process address space as the vulnerable app.
This attack works when: (1) the WebView has the Javascript-to-Java bridge enabled; and (2) the attacker is able to inject a malicious Javascript payload into the network traffic between the phone and the web server. 
Many apps are vulnerable to this type of attack because ad network libraries typically embed adverts into apps by displaying HTML and Javascript inside a WebView with the Javascript-to-Java bridge enabled. 
The fix for the addJavascriptInterface vulnerability breaks backwards compatibility, and therefore apps are only immune to this attack if their target API level is greater than 16 and they are running on an Android device with an OS version of 4.2 or greater.
Later analysis in Figure~\ref{fig:da_api} in this paper shows that 40\% of handsets connecting to the Google Play store are currently vulnerable to this attack.

\section{Data sources}
\label{sec:background}

We require two sources of data for this study: (1) information on the distribution of installed versions of Android over time across the Android ecosystem and (2) information on the critical vulnerabilities found to affect specific versions of Android.
These two datasets can then be combined to determine the proportion of handsets at risk of attack from specific vulnerabilities for particular periods of time.

\subsection{Versions of Android running on devices}

In our analysis we use historical data collected by the \da\ project~\cite{Wagner2013}.
Device Analyzer collects data from study participants who install the project Android app from the Google Play store.
Most study participants allow researchers around the world to access a subset of their device data, including the data presented in this paper.

The \da\ app collects a range of metadata from Android devices.\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.html}}
For this paper, we extracted the build string and API version from the device log each day.
The API version is a positive integer which increases when new features are added to the API.
Consequently security (bug) fixes do not result in a change in the API version.
The build string is a user-readable version string.
Fortunately most (\daOSVersionPercValidLines) entries in the data have a build string of the form `x.y.z opaque\_marker' and so it is possible to extract the Android version number `x.y.z'.
On a large proportion of devices `opaque\_marker' is a well defined build number\footnote{\url{https://source.android.com/source/build-numbers.html}} however different device manufacturers use different schema.

The \da\ project has collected data from \daNumDevices\ devices with a total of \daDeviceDays\footnote{Here we are only counting devices and days for which we have valid OS version data.} device days. The majority of devices only contribute data for a short period of time, however \daMonthsDevices\ devices have contributed data for more than \daMonths~months.

Where possible we indicate the uncertainty in our results by presenting them $\pm$ one standard deviation and give results to 3 significant figures, this occasionally results in `${} \pm 0$' when the standard deviation is less than the number of significant figures we are reporting.
This technique cannot account for systematic errors and therefore we explore the potential for systematic errors in the \da\ data in \S\ref{sec:representative}.

\subsection{Critical vulnerabilities}

We compiled a list of critical vulnerabilities in Android, containing information on the discovery and publication dates, the versions affected and which versions fixed the problem.
We only looked for critical vulnerabilities such as root vulnerabilities which did not require USB debugging to exploit.
Critical vulnerabilities allow a program to gain privileges equivalent in scope to root.
If an application exploits a critical vulnerability then it gains control of the device.
Some phones can be `rooted' by enabling USB debugging and using the special privileges of the ADB shell to root the device but only \daAdbEnabledPerc\dafoot\ of devices have USB debugging enabled.
This is not something that applications running on the phone can exploit to break out of the app sandbox and so we do not include those vulnerabilities.
Unfortunately, many published exploits use ADB for convenience and so determining whether the use of ADB is necessary to exploit the vulnerability can be difficult.

Some critical vulnerabilities are not traditional kernel vulnerabilities, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root equivalent privileges.
On some versions of Android (below version 4.1) malware could use known system-to-root escalation mechanisms but on all versions they have a greatly increased attack area for further privilege escalation and also have the ability to control all user internet traffic (via VPNs), brick the phone, remove and install apps, steal user credentials, read the screen and make as well as receive calls.
\todolater{break vulnerabilities down by attack vector?}
\avoTabAndVulns

We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, the AndroidVulnerabilities.org\footnote{\url{http://androidvulnerabilities.org/} \textbf{Note to reviewers:} we have removed our names and affiliation from the website for this review \textbf{with the exception of the contact page. Please do not visit the contact page until the review cycle has completed.}} (\avo) website.
We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
In addition, we have received submissions or amendments from \avoNumSubmitters\ individuals.
We collected data between \avoStartDate\ and \avoEndDate\ and so any information lost before the start of that period cannot be included.
\avo\ currently contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or device manufacturers.

Unless otherwise stated, we use \daNumVulnsUsed\ vulnerabilities in our analysis as shown in Table~\ref{tab:andvulns}.
We have chosen these vulnerabilities to fit the attack vectors introduced in \S\ref{sec:threatmodel}.
These vulnerabilities affect all Android devices regardless of manufacturer, and as a result our selected vulnerabilities will dominate security analysis of the Android ecosystem as a whole.
In many cases we could not match manufacturer- and model-specific vulnerabilities to individual devices in the \da\ data and therefore attempting to include device-specific vulnerabilities as well would introduce additional uncertainty in our results.
In contrast, with our chosen set of vulnerabiltiies, our analysis represents a lower-bound on the vulnerability of devices in the \da\ data set.

Tracking vulnerabilities is a manual task as they are not consistently recorded in other databases such as the CVE database.
In addition, the lack of a widely acknowledged unique identifier required manual analysis to identify whether two reports referenced the same vulnerability.
Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for critical Android vulnerabilities.
For some of the vulnerabilities without CVE numbers, Google confirmed that there was no CVE number and that they did not intend to get one, instead providing an Android bug number.

\subsection{Lifetime of a vulnerability}

The key events in the lifetime of a vulnerability do not always occur in the same order and are:\\
\textbf{creation} When a vulnerability was created in the source code.\\
\textbf{introducing release} When the first release was made containing the vulnerability.\\
\textbf{discovery} When the vulnerability is first discovered.\\
\textbf{exploit} When the vulnerability is first exploited.\\
\textbf{disclosure} When the vulnerability is first disclosed.\\ % publically or to a smaller set of people?
\textbf{fix} When a vulnerability was first fixed in the source code.\\
\textbf{fixing release} When the first release containing the fix was made (equivalent to `patch available'~\cite{Frei2010}).\\
\textbf{fix deployed} When the fix has been deployed to a sufficiently large proportion of the population that the vulnerability can be ignored (the ecosystem equivalent to the per instance `patch installed'~\cite{Frei2010}).
\todolater{Does this look nice?}

Establishing when a vulnerability starts to pose a threat to users is difficult.
Frei et al.~\cite{Frei2010} propose the definition of the {\bf time of disclosure}. This occurs when the information about the vulnerability is freely available to the public from a widely accepted and independent source and has been validated by security experts so that it has a risk rating.
Unfortunately before we collated this information much of it was not published by an independent source and lacked risk rating information, even months or years after they had been actively used.
Therefore this measure does not work.

Symantec's 2012 analysis of desktop malware has shown that after public disclosure, exploitation rates increase by 5 orders of magnitude~\cite{Bilge2012} and so from the point of view of widespread danger, the period between public \textbf{disclosure} and the date of \textbf{fix deployed} is the most critical.
However they also show that zero-day vulnerabilities are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
Hence when considering a vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those responding to CESG advice~\cite{CESG2013}, the critical period starts with \textbf{discovery} and continues until \textbf{fix deployed}.
Therefore, in our vulnerability calculations we prefer to use the earliest recorded date of \textbf{discovery}, even if that knowledge might have been confined to a particular device manufacturer or hobbyist.
%We cannot know if someone reported the vulnerability to the device manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they may know.

Unfortunately {\bf discovery} is the hardest point to obtain concrete data on as discovery may happen multiple times independently and not all discoverers will report their discovery.
{\bf Creation} and {\bf introducing release} are the earliest points which could be used, but the risk is mostly latent until someone discovers them and therefore we do not use these dates.
%, for example entities with high value targets may keep databases of unknown vulnerabilities for later use, in which case there is a high risk to potential targets from the never disclosed point of discovery.
%However if the discoverer never discloses the vulnerability to anyone or makes any use of it then there is little danger.
The date of first {\bf exploit} is a point when the risk is definitely high, but again a good adversary will not be detected when using such exploits.
The date of first {\bf fix} is, assuming that the fix is deliberate, a point at which the vulnerability is known at least within the organisation performing the fix and frequently implies an earlier discovery and notification by a third party; we determine this date from the Android Open Source Project repository by examining the authored-on date of the fixing commit.
%Once a {\bf fixing release} has been made then the vulnerability is widely known because it can be reverse engineered from the changes in the release~\cite{Brumley2008}.
%When the {\bf fix deployed} to a sufficiently large proportion of devices then the remaining risk is minimal.
Therefore, in this paper we use the earliest known date from \textbf{discovery}, \textbf{exploit}, \textbf{disclosure} or \textbf{fix}. A breakdown of the type of date used per vulnerability is shown in Table~\ref{tab:andvulns}.

\subsection{Distribution of vulnerabilities}
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/vulnerabilities_timeline}
 \caption{Timeline of vulnerabilities. For each vulnerability we show the dates of \textbf{discovery} and, when later, the dates of the first \textbf{fixing release}.}
 \label{fig:vulnerabilities_timeline}
\end{figure}

Figure~\ref{fig:vulnerabilities_timeline} shows the dates of \textbf{discovery} and, when later, the date of the first \textbf{fixing release}.
Some vulnerabilities (\emph{levitator}, \emph{zergRush}) were fixed in released versions of Android before they were discovered and so are shown as vertical lines, while others were known for months before a version of Android which fixed them was shipped.
During the period in which \da\ data was collected the date when a version of Android with the fix was observed on a \da\ device is taken as the date that version was released.
For vulnerabilities prior to the collection period we estimate the release date as best we can using publicly available data.
There is no canonical source of Android release dates, and our best guesses and supporting references are available from \avo.

The discovery dates of our vulnerabilities does not appear to be uniform. 
In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
The cause of this quiet period is unclear.
Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was no point in looking for new ones (from Figure~\ref{fig:proportioninsecure});
(ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and therefore attackers looking for vulnerabilities adjusted their focus.

@@

\section{Experiments}
\label{sec:results}
We present the results of our analysis showing that, on average, \daMeanInsecurityPerc\ of Android devices are exposed to critical vulnerabilities.
Devices, on average, apply \daUpdatesPerYear\ updates each year and \daPercUpdatesDowngrades\ of full version changes are downgrades to older versions.
Nexus devices are better than others, \emph{\daSecScoreBestmanufacturer} is the highest scoring device manufacturer and \emph{\daSecScoreBestoperator} is the highest scoring network operator.
%\todo{comparison of nexus update rates with iOS}
We describe four experiments.
We examine the vulnerability of the Android ecosystem as a whole (\S\ref{sec:exp:versionsecurity}, explore the upgrade cycle and vulnerability cycle of the ecosystem (\S\ref{sec:exp:android_ecosystem}), quantify the updates installed on particular devices (\S\ref{sec:exp:device_updates}) and compare the performance of different device manufacturers and network operators (\S\ref{sec:exp:security_score}).

%% In an experiment you test a hypothesis, it needs to be clear what the hypothesis is and what the conclusion of the test is

\subsection{Experiment 0: vulnerability of Android devices}\label{sec:exp:versionsecurity}

In the introduction, Figure~\ref{fig:proportioninsecure} described the proportion of Android devices susceptable to at least one critical vulnerability.

\subsubsection{Method} 
To investigate the behaviour of the Android ecosystem as a whole with respect to Android versions and vulnerabilities we used the OS version information from \da\ and the vulnerability data from \avo.
The \da\ data used here was collected between \daStartDate\ and \daEndDate.
The \avo\ data covers the period from \avoFirstDataDate\ to \avoLastDataDate.
For each device we collected daily version data  and recorded the running version and any associated vulnerabilities it was exposed to at that time (if any).
We then normalise these totals for each day by dividing through by the total number of devices with version information seen on that day.

\subsubsection{Results}

Following time from left to right in Figure~\ref{fig:proportioninsecure} we see all devices are initially \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011. This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
Once \emph{zergRush} was discovered in October 2011 the most devices are recorded as \emph{insecure} (red) as most devices were exposed to that vulnerability.
The remanding devices were already running a version of Android which fixed the \emph{zergRush} vulnerability and are therefore marked as  \emph{secure} (green).
From October 2011 until the discovery of \emph{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded. This means more and more devices are marked as \emph{secure} (green) because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received a minor update which may have contained a backported fix.
From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are exposed to known critical vulnerabilities.

\subsection{Experiment 1: behaviour of the Android ecosystem}\label{sec:exp:android_ecosystem}

\subsubsection{Method} As in Experiment 0, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.

\subsubsection{Results}
The proportion of devices in the \da\ data running different versions of Android each day is shown in Figure~\ref{fig:norm_os}.
It shows how old versions are gradually replaced by new ones, and the long tail of devices which do not see updates to more recent versions.

\begin{figure*}
 \centering
 \includegraphics[width=\textwidth]{figures/da_norm_os}
 \caption{Android versions in \da\ data over time. The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
 \label{fig:norm_os}
\end{figure*}

The vulnerabilities devices are exposed to are shown in Figure~\ref{fig:nvulnerabilities_heat}.
For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
The variation of the proportion of devices affected by a vulnerability with time tells us how badly a particular vulnerability affected the Android platform.
In July 2011 at the beginning of the \da\ data the \emph{exploid}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}} and \emph{levitator}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/levitator}} vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices (\daVulnAPKDuplicateFileOctoberPerc) were still vulnerable.
%\begin{figure}%[!b]
%\centering
%\includegraphics[width=\columnwidth]{figures/vulnerabilities}
%\caption{Proportion of devices exposed to each vulnerability with time (1.0 is 100\%)}
%\label{fig:vulnerabilities}
%\end{figure}

In 2013 three vulnerabilities were found in the way which Android verified the signatures on APKs.
These allowed the creation of malicious APKs which appear to be signed as system APKs -- which have root equivalent privileges.
Figure~\ref{fig:nvulnerabilities_heat} shows how the the \emph{APK signing vulnerabilities} affected all devices and took months to get fixed for any device.
However what is perhaps more worrying is the long tail on the \emph{Gingerbreak}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}}, \emph{levitator}, \emph{exploid} and \emph{zergRush}\footnote{\url{http://androidvulnerabilities.org/vulnerabilities/zergRush}} vulnerabilities which are more dangerous root vulnerabilities (not requiring new APK installation) and which still affect a significant proportion of devices years later.

\begin{figure*}
 \includegraphics[width=\textwidth]{figures/nvulnerabilities_heat.pdf}
 \caption{Proportion of devices affected by different vulnerabilities. The prefix `(maybe)' indicates `maybe secure' and `(definite)' indicates `definitely insecure'. The first few vulnerabilities are all `maybe secure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices. The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}}
 \label{fig:nvulnerabilities_heat}
\end{figure*}

\subsection{Experiment 2: Updates to particular devices}\label{sec:exp:device_updates}
Those graphs summarise data across all the devices, however one of the advantages of the \da\ data is that it allows us to look at what happens to individual devices over time.

\subsubsection{Method}
As in Experiment 0, we used the version information for each device to calculate which critical vulnerabilities each device was susceptible to on a daily basis.
We also recorded when the version changed and which versions it changed from and to.
In the \da\ data we have over 1000 devices contributing data in any particular week.
However most devices only contribute for a short period of time and so we do not observe updates happening on every device.
Instead we have a hopefully representative sample of updates which happened while \da\ was installed on the devices.
\da\ cannot distinguish between a device being replaced and the \da\ app being removed as while multiple devices can be linked to the same user account, that is a manual process and few users do it.

\subsubsection{Results}
\begin{figure}
 \centering
 \begin{subfigure}[b]{\columnwidth}
  \includegraphics[width=\columnwidth]{figures/device-data-all-os}
  \caption{OS versions over time. Black vertical lines indicate that the build number changed without changing the OS version at that point.}
  \label{fig:device_data_os}
 \end{subfigure}
 \begin{subfigure}[b]{\columnwidth}
  \includegraphics[width=\columnwidth]{figures/device-data-all-security}
  \caption{Number of vulnerabilities affecting each device over time}
  \label{fig:device_data_security}
 \end{subfigure}
 \caption{The top \daNumDeviceDataDevices\ devices by days of contribution in the \da\ data. One strip per device handset.}
\end{figure}
The longitudinal data on the number of vulnerabilities affecting the \daNumDeviceDataDevices\ devices which have contributed the most days of data to \da\ changes over time is shown in Figure~\ref{fig:device_data_security}.
The trend that we saw in Figure~\ref{fig:proportioninsecure} of security improving and then getting worse is also shown here.
It shows how some devices had vulnerabilities, which were fixed, and then further vulnerabilities were discovered, and for these devices, mostly not fixed.
This implies that these devices had been abandoned by the device manufacturer and were not receiving updates which is confirmed by Figure~\ref{fig:device_data_os} which shows which OS versions those devices were running.
Some devices start off in 2011 exposed to known security vulnerabilities and are still exposed to those and additional ones in 2014.

\begin{figure}
 \centering
 \includegraphics[width=0.6\columnwidth]{figures/from_to_updates.pdf}
 \caption{Updates between different Android versions in the \da\ data}
 \label{fig:from_to_updates}
\end{figure}
\begin{figure}[h]
 \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
 \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities}
 \label{fig:weekly_security_updates}
\end{figure}
We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
Mostly the dark cells are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgradesNominal) which skip more than \daNumUpdatesSkippedBig\ versions.
Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngradesNominal) when older versions of Android are installed on to devices.
Possible reasons why users are downgrading are to free up space on their device, to make it easier to root or because a new version introduced bugs.

The number of devices getting security updates each week, is shown in Figure~\ref{fig:weekly_security_updates}.
\begin{itemize}
 \item Updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities are shown in red.
 \item Updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability are shown in yellow.
\item Updates which did not fix security vulnerabilities (because there were no known security vulnerabilities in the version of Android they were already running) are shown in green.

\end{itemize}

By taking the number of updates observed and the number of device days of data we have collected we can compute the number of updates received by a device per year to be \daUpdatesPerYear.
This compares badly with the number of critical vulnerabilities discovered per year of between \avoVulnsPerYearAllAndroid\ (affecting all Android) and \avoVulnsPerYear\ (including the device manufacturer specific ones in \avo).

@@

\section{Threats to validity}
\label{sec:validity}
\subsection{The \da\ data gives a conservative estimate of the Android version distribution}
\label{sec:representative}
\begin{figure}
 \centering
 \begin{subfigure}[b]{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/googleplayapi}
 \caption{Google Play data on proportion of devices running different Android API versions}
 \label{fig:play_api}
\end{subfigure}
\begin{subfigure}[b]{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
 \caption{\da\ data on proportion of devices running different Android API versions}
 \label{fig:da_api}
\end{subfigure}
\caption{Monthly Android API version data}
\end{figure}
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
 \caption{Difference between \da\ and Google Play data on the proportion of devices running different Android API versions}
 \label{fig:da_gp_comp_diff}
\end{figure}
The data from \da\ we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
However Google has published API version information almost every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions is similar to the Google Data on API versions then the \da\ data on OS versions should be representative.
Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
Figure~\ref{fig:da_gp_comp_diff} shows the difference between Figures~\ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
It shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
\todolater{we want a statistical metric to claim this strongly with.}
This allows us to have confidence in the OS version information.

\subsection{Changes in the \da\ sample}
\label{sec:da_changes}
The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app on Google Play as there has been no advertising and few other attempts to increase usage.
However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular areas and network operators.
One study was conducted in Norway with 654 installs of \da\ and one was conducted in Bangladesh where 2463 users installed \da.
This is responsible for the sudden change in distribution of OS versions in \da\ shown towards the end of Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
%We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
To investigate the sensitivity of our results to the Bangladesh study we truncated our data before this study started and found the percentage of vulnerable devices to be 87.7\% rather than \daMeanInsecurityPercNominal, and the percentage of devices running the latest version to be 5.43\% rather than \daUpdatednessPercNominal.
Hence removing this anomaly would have no effect on our conclusions.
Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.

>>>>

\documentclass{sig-alternate}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[space]{grffile}
\graphicspath{figures/}
\usepackage{import}
\usepackage{url}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{siunitx}
\usepackage{todonotes}%[disable]
\let\OldTodo\todo
\renewcommand{\todo}{\OldTodo[inline]}
\newcommand{\todolater}[1]{\todo{#1}}%{}%
\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
\addbibresource{securityupdates.bib}
\renewcommand{\bibfont}{\small}

@@

@@

% Blinding function
\newcommand{\identifying}[1]{#1}%{}%
\newcommand{\blindauthors}[1]{#1}%{Paper \#XX}%

@@

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{CCS}{2015 Denver, Colorado, USA}
%\CopyrightYear{2015} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{TODO}

\numberofauthors{3}
\author{
Daniel~R.~Thomas
\and
Alastair~R.~Beresford\\
       \affaddr{Computer Laboratory}\\
       \affaddr{University of Cambridge}\\
       \affaddr{Cambridge, United Kingdom}\\
       \email{Firstname.Lastname@cl.cam.ac.uk}
\and
Andrew~Rice
}

@@

\begin{abstract}
We find there is significant variability in the timely delivery of security updates across different device manufacturers and mobile network operators, and we define a security metric to rank the performance of device manufacturers and network operators in our data set.
TODO
\end{abstract}

% A category with the (minimum) three required fields
\category{}{}{}

\terms{}

\keywords{}

\section{Introduction}

@@

@@

In summary, the contributions of this paper are:
\begin{itemize}
 \item We measure the security of Android according to three metrics and compare different device manufacturers and device models to allow device purchasers to differentiate between them based on security.
\end{itemize}

@@

@@

\section{Related work}
The update process for apps, security fixes and OS upgrades also needs to be secure.
Unfortunately, package management systems designed to provide secure updates have been found to contain vulnerabilities~\cite{Cappos2008} and many software update systems fail to authenticate the connection between the device and the update server or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
Android does authenticate update binaries and Google Play downloads them over a secure connection~\cite{Viennot2014}.
In this paper we have analysed four critical vulnerabilities in the Android app update mechanism: APK unsigned shorts, APK unchecked name, APK duplicate file and Fake ID.
Other work has demonstrate complex and subtle errors exist in the Android app update process.
For example, the process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.
